Network based multifactorial modelling microRNA-target interaction.

### 2019-09-01 article.Rmd
git problem
Article edit

MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key regulator of gene expression through various post-transcriptional mechanisms. Although the mechanisms by which miRNA represses are not fully understood, miRNAs predominantly repress their targets. Repressive activities of miRNAs vary depending on many factors that are significant to miRNA:target interactions. These factors include miRNA:target binding energy, binding location in target sequence, base pairing types between miRNA and target, abundance of miRNAs and targets [@grimson_microrna_2007]. Binding energies of miRNA:target complexes vary based on nucleotide context and determine folding stability of complex [@cao_predicting_2012]. It has been demonstrated that the binding energy between miRNA and target indicates stability or affinity of complex [@helwak_mapping_2013] and does not directly determine repressive activity of miRNA [@cao_predicting_2012]. Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5’end bind to specific sequence located in 3’UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005]. In recent studies, it has been shown that miRNAs can interact with targets via sequences located in regions such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013; @moore_mirnatarget_2015]. These studies also showed that binding location could indicate functionality of miRNA:target interaction or be effective on abundance of targets(TOFIX abundance etkisi derken, bazen azaltabilir, bazen arttırabilir, manasında mı? answer: evet hocam anlaşılmıyorsa değiştirebilirim). It has been shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence that is perfectly complementary with targets, seed region at the 5’ end of miRNAs [@bartel_micrornas:_2009; @grimson_microrna_2007]. On the other hand, some researchers have reported that seed sequence of miRNA can have mismatches or bulged/wobble nucleotides and may locate in region other than 5'end of miRNAs [@hafner_transcriptome-wide_2010; @helwak_mapping_2013]. On top of all these factors, abundance of miRNAs and targets and miRNA:target ratio in cells predominantly affect efficiency of miRNA:target interaction [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].

@@

[#add]Similar with these approach, Silveira et al. have developed miRmapper package for using with R programme [@da2018mirmapper]. Researchers have utilized the adjacency matrix to associate miRNAs using differentially expressed genes and found similarities of miRNAs with this approach. Authors who utilized the approach in dataset of bladder cancer cell line have identified importance of miRNAs in network, and miRNAs which work synergistically.

>>>>

MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key regulator of gene expression through various post-transcriptional mechanisms. Although the mechanisms by which miRNA represses are not fully understood, miRNAs predominantly repress their targets. Repressive activities of miRNAs vary depending on many factors that are significant to miRNA:target interactions. These factors include miRNA:target binding energy, binding location in target sequence, base pairing types between miRNA and target, abundance of miRNAs and targets [@grimson_microrna_2007]. Binding energies of miRNA:target complexes vary based on nucleotide context and determine folding stability of complex [@cao_predicting_2012]. It has been demonstrated that the binding energy between miRNA and target indicates stability or affinity of complex [@helwak_mapping_2013] and does not directly determine repressive activity of miRNA [@cao_predicting_2012]. Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5’end bind to specific sequence located in 3’UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005]. In recent studies, it has been shown that miRNAs can interact with targets via sequences located in regions such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013; @moore_mirnatarget_2015]. These studies also showed that binding location could indicate functionality of miRNA:target interaction or be effective on abundance of targets(TOFIX abundance etkisi derken, bazen azaltabilir, bazen arttırabilir, manasında mı? answer: evet hocam anlaşılmıyorsa değiştirebilirim). It has been shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence that is perfectly complementary with targets, seed region at the 5’ end of miRNAs [@bartel_micrornas:_2009; @grimson_microrna_2007]. On the other hand, some researchers have reported that seed sequence of miRNA can have mismatches or bulged/wobble nucleotides [@chi2012alternative] and may locate in region other than 5'end of miRNAs [@hafner_transcriptome-wide_2010; @helwak_mapping_2013]. On top of all these factors, abundance of miRNAs and targets and miRNA:target ratio in cells predominantly affect efficiency of miRNA:target interaction [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].

@@

[#Add]Similar with this approach, Silveira et al. have developed miRmapper package for using with R programme [@da2018mirmapper]. Researchers have utilized the adjacency matrix to associate miRNAs using differentially expressed genes and found similarities of miRNAs with this approach. Authors who utilized the approach in dataset of bladder cancer cell line have identified importance of miRNAs in network, and miRNAs which work synergistically.

[#Add] Today, there are many of tools that provide predicted or experimentally validated miRNA:target interaction dataset. The datasets that contain weak, proteomic analysis like western blot, and strong, crosslinking and immunoprecipitation CLiP, experimentally evidence are more reliable sources than prediction datasets. While in experimentally weak supported datasets it is not known that indirect or direct target of miRNAs in cell, high-throughput methods ensure handling of bounded miRNA:target pairs. For this reason we have preferred to utilize high-throughput datasets in our application. We collected datasets that depend on chimeric [@helwak_mapping_2013; @moore_mirnatarget_2015] reading of miRNA: target pairs from among these datasets because they contain exact complementary seed regions. Although these experimental sources include exact binding sites, they do not provides which miRNA is functional or not [@liu2019prediction]. We have used the features of miRNA:target pairs and scaled these parameters inside groups that is consisted according to ,ceRNA (Competing Endogenous RNA), genes targeted by the same miRNA. *In scaling process, we have considered previous studies. BURAYA O ÇALIŞMALARI ATIFLA*
